\begin{tabulary}{\textwidth}{L|J}
    
    \toprule
    \textbf{Action}                 & \textbf{Key}              
    \\ \midrule
    
    \multicolumn{2}{c}{\textit{Setting selection.}}
    \\ \midrule

    Add a spot or link to the selection.  & \keys{\shift+Left-click}. Do it again to remove a selected spot or link from the selection.
    \\ \midrule
    
    Select all spots and links. & \keys{\ctrl+A}
    \\ \midrule
    
    Select just all the spots. & \keys{\ctrl+\Alt+A}
    \\ \midrule
    
    Select just all the links. & \keys{\ctrl+\shift+A}
    \\ \midrule
    
    Select all the spots and links in the track of the currently focused spot. & \keys{\shift+Space}
    \\ \midrule
    
    Select all the spots and links \textit{upward} (backward in time) in the track of the currently focused spot. & \keys{\shift+Page\ \arrowkeyup}
    \\ \midrule
    
    Select all the spots and links \textit{downward} (forward in time) in the track of the currently focused spot. & \keys{\shift+Page\ \arrowkeydown}
    \\ \midrule
    
    Add the previous / next spot in time to the selection.      & \keys{\shift + \arrowkeyup} / \keys{\shift + \arrowkeydown}
    \\ \midrule

    Add all spots to the selection from current spot to the beginning of a branch.  & \keys{\shift + \Alt + \arrowkeyup}
    \\ \midrule
    
    Add all spots to the selection from current spot to the end of a branch.  & \keys{\shift + \Alt + \arrowkeydown}
    \\ \midrule

    Add all spots to the selection from current spot to the beginning of another branch. & \keys{\shift + \ctrl + \Alt + \arrowkeyup}
    \\ \midrule

    Add all spots to the selection from current spot to the end of another branch.  & \keys{\shift + \ctrl + \Alt + \arrowkeydown}
    \\ \midrule

    \multicolumn{2}{c}{\textit{Selection in TrackScheme.}}
    \\ \midrule
    
    Add the spot to the left / right of the focus to the selection.  & \keys{\shift+\arrowkeyleft} / \keys{\shift+\arrowkeyright}
    \\ \midrule
    
    Draw  a selection box   & \keys{Left-click} and \keys{Drag}.
    \\ \bottomrule
  
\end{tabulary}
